---
title: "Analyzing Youth Voter Turnout Through the Lens of Descriptive Representation"
author: "Matt Lehman"
date: "`r format(Sys.time(), '%b %d, %Y') `"
output: 
  html_document: 
    fig_height: 4
    highlight: pygments
    theme: spacelab
---

## Setup and Data Manipulation

### Setup
Load packages
```{r pkgs, messages = FALSE}
library(dplyr)
library(feather)
library(lubridate) 
library(ggplot2)
library(stargazer)
```

### Obtain Data

#### Voter File Data
Connect to the database and read in the table
```{r DB conn}
path <- "my_data.feather"

if (!file.exists(path)) {
  library(DBI)
  library(RSQLite)
  con <- dbConnect(RSQLite::SQLite(), dbname = "../data/govt990.db") # connect to DB
  # Send the table to feather so I can access it in Python and R
  write_feather( dbGetQuery(con, 'SELECT * FROM nys_young_voters;'), path )
  
  dbDisconnect(con) # disconnect from DB
  warning("I should really just do all of this in the Python block")
}
```

For real, sticking to R for the rest of the analysis.

```{r data to R}
# Get data in R dataframe
df <- read_feather(path)
```

#### Census Data
Get data from the 2010 and 2014 Census. I got the files from https://www.nhgis.org/. I downloaded files with
descriptive headers, so I will skip reading the first row.
```{r read census data}
census.cd.10 <- read.csv("census_data/nhgis0037_ds172_2010_cd110th-112th.csv", skip = 1)
census.cd.14 <- read.csv("census_data/nhgis0036_ds215_20155_2015_cd113th-114th.csv", skip = 1)
```

I need to manipulate the Census data in a few ways. First, I only need to data from New York. 
Also, the Census supplied age data by sex. I am going to manipulate the dataframe to have the total
number of people aged 18 -- 24 living in each Congressional District.
```{r manipulate census data}
young_pop10 <- census.cd.10 %>%
  filter(State.Name == "New York") %>%
  group_by(Congressional.District..111th.Congress..Code) %>%
  summarise(young_population = 
              sum(Male..18.and.19.years,
                  Male..20.years,
                  Male..21.years,
                  Male..22.to.24.years,
                  Female..18.and.19.years,
                  Female..20.years,
                  Female..21.years,
                  Female..22.to.24.years))
names(young_pop10)[1] <- "CD"

young_pop14 <- census.cd.14 %>%
  filter(State.Name == "New York") %>%
  group_by(Congressional.District..2013.2017..113th.114th.Congress..Code) %>%
  summarise(young_population = 
              sum(Estimates..Male..18.and.19.years,
                  Estimates..Male..20.years,
                  Estimates..Male..21.years,
                  Estimates..Male..22.to.24.years,
                  Estimates..Female..18.and.19.years,
                  Estimates..Female..20.years,
                  Estimates..Female..21.years,
                  Estimates..Female..22.to.24.years))
names(young_pop14)[1] <- "CD"
```


#### Turnout Rates

Get voter turnout rates. First, just for 2014 

```{r Turnout rates}
# get voters who were young and voted in 2014, 
# then filter out duplicate voter IDs

# It is sorted to filter out older registration records.
# TODO: a better way to do this would be to filter out duplicated
# records with the constaint that the final record has to be before the election
vote14 <- df %>%
  filter(Young2014 & Voted2014) %>%
  arrange(desc(Reg_Date))

# dup <- vote14[duplicated(vote14$StateID), ] %>% 
#   select(Reg_Date, StateID) %>%
#   arrange(desc(Reg_Date), StateID)

full <- nrow(vote14)
vote14 <- vote14[!duplicated(vote14$StateID), ]
warning( paste("Elimiated", full - nrow(vote14), "rows through de-duplication for 2014 data", sep = " ") )
# Aggregate voter turnout by Congressional District
turnout14 <- vote14 %>%
  group_by(CD) %>% 
  summarise(turnout = sum(Voted2014))

# Match to Census population by Congressional district to get rate
turnout14 <- turnout14 %>% 
  left_join(young_pop14) %>%
  mutate(turnout_rate = turnout / young_population)

# hist(turnout14$turnout_rate) # That looks really low
summary(turnout14$turnout_rate) # There are some *very* low values, and max turnout is ~ 8%
sd(turnout14$turnout_rate)

# Now do this for 2010 (did 2014 first because data are better)
vote10 <- df %>% 
  filter(Young2010 & Voted2010) %>% 
  arrange(desc(Reg_Date))
full <- nrow(vote10)
vote10 <- vote10[!duplicated(vote10$StateID), ]
warning( paste("Elimiated", full - nrow(vote10), "rows through de-duplication for 2010 data", sep = " ") )

# Aggregate voter turnout by Congressional District
# Match to Census population by Congressional district to get rate
turnout10 <- vote10 %>%
  group_by(CD) %>% 
  summarise(turnout = sum(Voted2010)) %>% 
  left_join(young_pop10) %>% 
  mutate(turnout_rate = turnout / young_population)


# hist(turnout10$turnout_rate) # That looks really low but a bit better
summary(turnout10$turnout_rate) # There are some *very* low values, and max turnout is ~ 15%
sd(turnout10$turnout_rate)

turnout10$year <- 2010
turnout14$year <- 2014

turnout <- rbind(turnout10, turnout14)
```
The turnout rates look very low, but for now I will continue with the analysis. I do want to check if there
were districts that seem to have articificially lower turnout rates. I will consider removing those 
districts from the analysis until I have better data.

```{r map turnout, eval = FALSE}
# Map turnout to see if anything looks "weird"
# I can't install leaflet, so this will have to wait. 
```

#### Candidate Data
```{r read candidate data}
candidates <- read.csv("candidate_data/nys_candidates_ages.csv")

# candidates$election_districts <-gsub("\\{|\\}", "",candidates$election_districts)

#' Return the last element in a list
#'
#' This function will be used to return the last
#' district that a candidate ran in, which
#' should be the district they ran in during 2014.  
get_last <- function(array) {
  # Array is a list
  return( as.integer( array[length(array)] ) )
}

# For 2010, assume the second to last district (if more than one) --> district at last census usually, if there was a change
get_secondlast <- function(array) {
  # Array is a list
  filtered <- unique(array)
  if (length(filtered == 1)) {
    return( as.integer(filtered[1]) )
  }
  else {
    return( as.integer(array[length(filtered) - 1]) )
  }
}

# Recode candidate status
candidates <- candidates %>% 
  mutate(sta10 = recode(sta10, C = 1, O = 2, I = 3, .default = 0),
           sta14 = recode(sta14, C = 1, O = 2, I = 3, .default = 0))

# Filter only candidates who ran in 2014 and did not run unopposed 
candidates.14 <- candidates %>%
  filter(gen14 == 1)

# For 2014, assume the last district
election.districts <- strsplit(as.character(candidates.14$election_districts), ",")
candidates.14$district <- sapply(lapply(election.districts, function(x) get_last(x)), function(x) return(x[1]))

# Get candidate age for 2014 midterm
candidates.14$age <- as.period(interval(as.Date(candidates.14$dob, format = "%m/%d/%Y"), "2014-11-4"), units = "year")$year

# For each district, get the youngest age and the average age
districts.14 <- candidates.14 %>% 
  group_by(district) %>%  
  summarise(youngest_age = min(age, na.rm = T),
            average_age = mean(age, na.rm = T),
            status = max(sta14))

warning("\nNo age data for district 4 in 2014, filtering out.")
districts.14 <- districts.14 %>%
  filter(district != 4) %>%
  filter(!is.na(district))

# Do the same thing for 2010
# Filter only candidates who ran in 2010 and did not run unopposed 
candidates.10 <- candidates %>%
  filter(gen10 == 1)

# For 2010, assume the second to last district (if more than one) --> district at last census usually, if there was a change
election.districts <- strsplit(as.character(candidates.10$election_districts), ",")
candidates.10$district <- sapply(sapply(election.districts, function(x) get_secondlast(x)), function(x) return(x[1]))
# Get candidate age for 2014 midterm
candidates.10$age <- as.period(interval(as.Date(candidates.10$dob, format = "%m/%d/%Y"), "2010-11-2"), units = "year")$year

# For each district, get the youngest age and the average age
districts.10 <- candidates.10 %>% 
  group_by(district) %>%  
  summarise(youngest_age = min(age, na.rm = T),
            average_age = mean(age, na.rm = T),
            status = max(sta10))

# Filter out NA
districts.10 <- districts.10 %>%
  filter(!is.na(district))
```

#### Master Dataframe

Create one dataframe to link all of the district-level data together

```{r join all data}
all14 <- turnout14 %>% 
  inner_join(districts.14, by = c("CD" = "district")) 

all10 <- turnout10 %>% 
  inner_join(districts.10, by = c("CD" = "district"))

master <- rbind(all10, all14) %>% arrange(CD, year)
```

```{r make age and year categorical}
master$age_cat <- "older" # init as older
master$age_cat[master$youngest_age <= 30] <- "young" 
master$age_cat <- factor(master$age_cat, levels = c("older", "young"))
master$year <- factor(master$year, levels = c(2010, 2014))
master$CD <- as.factor(master$CD)

master <- master %>% 
  mutate(inc_status = factor(case_when(status == 1 ~ "open",
                                TRUE ~ "incumbent")))
```


## Analysis
### One-Way Comparisons

```{r one-way plots}
fig.num <- 0
my_pal <- c("#FF6347" ,"#104E8B") # define my color palette

# Plot histogram of youth turnout
fig.num <- fig.num + 1
youth_turnout <- ggplot(master) + geom_histogram(aes(x = turnout_rate), bins = 8,
                                                 fill = my_pal[2]) +
  labs(title = "Youth Voter Turnout", x = "Turnout Rate", y = "Count", 
       caption = paste("Figure ", fig.num, ": Histogram of youth voter turnout.", sep = ""))
youth_turnout
```

### Two-Way Comparison
```{r two-ways plots}
# Plot turnout by age, split plots by year, and color by age category
# First, do continous. 

fig.num <- fig.num + 1
compare_plot <- ggplot(master %>% filter(year == 2014)) + 
  geom_point(aes(x = youngest_age, y = turnout_rate, col = age_cat), size = 2) +
  labs(title = "Youth Voting", x = "Youngest Candidate's Age", y = "Turnout Rate",
       caption = paste("Figure ", fig.num, ": Youth voter turnout by youngest candidate's age.", sep = "")) +
  scale_color_manual(values = my_pal, name = "Youngest \nCandidate",labels = c("Older", "Young"), guide = guide_legend(reverse=TRUE)) +
  theme_bw()
compare_plot

# fig.num <- fig.num + 1
# inc_plot <- ggplot(master) + geom_point(aes(x = youngest_age, y = turnout_rate, col = inc_status), size = 2) +
#   facet_grid(. ~ year) +
#   labs(title = "Youth Voting", x = "Youngest Candidate's Age", y = "Turnout Rate",
#        caption = paste("Figure ", fig.num, ": Youth voter turnout by youngest candidate's age and seat status.", sep = "")) +
#   scale_color_manual(values = my_pal, name = "Candidate \nStatus",labels = c("Incumbent", "Open"), guide = guide_legend(reverse=TRUE)) +
#   theme_bw()
# inc_plot

# It doesn't make sense to inculde an incumbency effect since only two seats a year
# were open. 
```

```{r two way tables}
master %>% 
  group_by(age_cat, year) %>% 
  summarise(avg = mean(turnout_rate), count = n()) %>% 
  arrange(year) # %>% xtable::xtable()
```


The continous graph shows turnout seems to increase with higher age in 2010 and decrease with higher age in 2014. 

The binned graph shows that younger candidates were associcated with the higher turnout rates in 2014. It's also important
to note there were only two youngest candidates 35 and under in 2010 and five in 2014.  

I am not trying to prove the the older a candidate gets, the lower the youth turnout rate. My hypothesis is that 
younger candidates (on a absolute scale) raise turnout relative to other candidates. Therefore, the categorical age
variable is the correct measure. 

### Logistic Regression

I am going to start with a logistic regression model predicting percentage of young people who vote in a district. 
```{r}
logit <- glm(data = master %>% filter(year == 2014), cbind(turnout, young_population - turnout) ~  inc_status + age_cat , family = "binomial")
summary(logit)

# Random effect for year
# library(lme4)
# logit.mixed <- glmer(data = master, cbind(turnout, young_population - turnout) ~ (1 | year) + inc_status + age_cat, family = "binomial")
# summary(logit.mixed)

# Random effect for Congressional District -- makes no sense since CD boundaries change
# Random effect for some geographic boundary might be a good idea
# logit.mixed2 <- glmer(data = master, cbind(turnout, young_population - turnout) ~ (1 | year) + (1 | CD) + age_cat, family = "binomial")
# summary(logit.mixed2)
```

### Visualize Model
```{r}
ages <- as.factor(c("young","older"))
stati <- as.factor(c("incumbent", "open"))
X <- expand.grid(data.frame(age_cat = ages, inc_status = stati))
preds <- predict.glm(logit, newdata = X,  type = "link", se.fit = T)
everything <- cbind(X, preds$fit, preds$se.fit)
names(everything) <- c("age_cat", "inc_status", "fit", "se")
# Make 95% CI
everything <- everything %>% 
  mutate(lower = fit -  se * qnorm(.975),
         upper = fit + se * qnorm(.975)) %>% 
  # Now change log odds to probabilities
  mutate(lower = exp(lower) / (1 + exp(lower)),
         fit = exp(fit) / (1 + exp(fit)),
         upper = exp(upper) / (1 + exp(upper)))
 
fig.num <- fig.num + 1 
pred_plot <- ggplot(everything, aes(x = as.factor(inc_status))) + 
  geom_linerange(aes(ymin = lower, ymax = upper, col = age_cat), size =1.21, position = position_dodge(.2)) +
  geom_point(aes(y = fit, col = age_cat), 
                  position = position_dodge(.2), size = 4) +
  labs(title = "Model Predictions (95% Interval)", x = "Seat Status", y = "Turnout Rate",
        caption = paste("Figure ", fig.num, ": Predictions of youth voter turnout using the logistic model.", sep = "")) +
  scale_color_manual(values = my_pal, name = "Youngest \nCandidate",labels = c("Older", "Young"), guide = guide_legend(reverse=TRUE)) +
  scale_x_discrete(labels = c("Incumbent", "Open")) +
  theme_bw()
pred_plot
```


#### Make Presentation
```{r powerpoint, eval = F}
library(officer)
library(magrittr) # `magrittr` supposedly makes officer usage easier

# Make slides with images
slides <- read_pptx() %>% 
  # Compare turnout by candidate age
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with_gg(youth_turnout) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with_gg(compare_plot) %>% 
  add_slide(layout = "Title and Content", master = "Office Theme") %>% 
  ph_with_gg(pred_plot)
print(slides, target = "images.pptx") 
```

#### Output Table
```{r, eval = F}
stargazer(logit, out = "logit.html", type = "html",
          title = "Logistic Regression",
          covariate.labels = c("Intercept", "Open Seat", "Young Candidate"),
          dep.var.caption = "Youth Voter Turnout Rate",
          dep.var.labels.include = F,
          ci = T, intercept.bottom = FALSE,
           omit.stat = c("aic", "bic", "LL"))
```
### Nominial Age Categories
Run the analysis again, but this time break age into three different
categories.
```{r three age categories}
master <- master %>% 
  mutate(age_cat3 = case_when(
    youngest_age <= 30 ~ "Young",
    youngest_age <= 50 ~ "Mid-Aged",
    youngest_age <= 100 ~ "Old"
  ))
logit.3 <- glm(data = master %>% filter(year == 2014), cbind(turnout, young_population - turnout) ~  inc_status + age_cat3, family = "binomial")

summary(logit.3)
# stargazer(logit.3, out = "logit_3.html", type = "html",
#           title = "Logistic Regression with Three Levels",
#           covariate.labels = c("Intercept", "Open Seat", "Young Candidate", "Old Candidate"),
#           dep.var.caption = "Youth Voter Turnout Rate",
#           dep.var.labels.include = F,
#           ci = T, intercept.bottom = FALSE,
#            omit.stat = c("aic", "bic", "LL"))

```

