---
title: "Explore NY Data"
author: "Matt Lehman (mhl84)"
date: "March 26, 2018"
output: 
  html_document:
    toc: true
    theme: readable
    highlight: haddock
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages, echo = FALSE, message=FALSE}
library(data.table)
library(dplyr)
```

### Read and Clean Candidate Data
```{r}
# data from https://www.fec.gov/data/candidates/?office=H&state=NY&has_raised_funds=true
# codebook for a similar dataset here: https://classic.fec.gov/finance/disclosure/metadata/DataDictionaryCandidateMaster.shtml
candidates <- fread("candidates-2018-03-26T12_17_18.csv")

# Let's get only candidates who ran in 2010, or 2014
filter.years <- c("2010", "2014")

# Remove the braces from the cycles field 
candidates$election_years <-gsub("\\{|\\}", "",candidates$election_years)
# Make a list. Each entry is a vector of the election years a canidate ran in
elections.list <- strsplit(candidates$election_years, ",")

# list apply to find if the cycles contain the years we want
# and then sapply to sum the resulting vectors of T/F per candidate
# End result is 1 if candidate ran in that year and 0 if not
# candidates$run2006 <- sapply(lapply(elections.list,function(x) "2006" %in% x), function(x) sum(x))
candidates$run2010 <- sapply(lapply(elections.list,function(x) "2010" %in% x), function(x) sum(x))
candidates$run2014 <- sapply(lapply(elections.list,function(x) "2014" %in% x), function(x) sum(x))

# Filter only candidate who ran in 2010 or 2014
candidates.filtered <- candidates %>%
  filter(run2010 + run2014 > 0)
```

```{r export inital file, eval = F}
# Export to .csv so I can add DOB
cols <- c("name", "office", "party", "election_districts", 
          "election_years", "cycles", "candidate_status", "candidate_id") # cols to export
write.csv(candidates.filtered %>% select(cols), "nys_candidates.csv")
```

I wanted to narrow down the dataset to only candidates that ran for open seats
so I would not have to control for incumbency advanatages. 
```{r, eval=FALSE}
# Make a new dataframe for those candidates who ran for open seats
candidates.no.incumbent <- candidates.filtered %>%
  filter(incumbent_challenge == "O")
# That's only 35 races
# It might make the project easier to look at only a few elections and have those
# elections be a simple case (ie won't have to control for incumbency factors)
# How many elections were actually in this filtered set?
length(unique(candidates.no.incumbent$district))
```
This was too small of a subsample. The effects I'm interested in should be apparent in the full dataset anyway. 

##### What I need to do for candidates:

- Get candidates' age at time of election

- Filter to only candidates who ran in the general election

- Consider filtering out third party candidates

###### What I need to do for voters:

- Find out how many voters by age group voted in each midterm election

- Validate counts against aggregated totals in Census files (look at the Pomante paper)

- Get VEP by Congressional district and use that to get turnout rates